Forecasting interrupted time series

### 2024-05-16 response_letter/response_letter_JORS.qmd

---
title: "Response letter - Manuscript ID 249381018"
bibliography: bibliography.bib
format:
  pdf:
    pdf-engine: pdflatex
    papersize: a4
    geometry: "margin = 2.5cm"
    include-in-header: header.tex
---

@@

# Editor comment

_Your manuscript entitled "Forecasting interrupted time series", which you submitted to Journal of the Operational Research Society, has been reviewed. The referee comments are included at the bottom of this letter. The referee(s) would like to see some revisions made to your manuscript before publication. Therefore, I invite you to respond to the referee(s)' comments and revise your manuscript. When you revise your manuscript please highlight the changes you make in the manuscript by using the track changes mode in MS Word or by using bold or coloured text._

**Response:** Thank you for giving us an opportunity to revise the manuscript. We have thoroughly reviewed the feedback provided by the referees, and made some changes as a result. We have highlighted all changes with coloured text in both the response letter and the manuscript.

We also wanted to provide further clarification to the editor regarding the comment provided by Reviewer 4.

- Reviewer 4 (comment 2) suggested including mathematical equations for each strategy. While we could add these equations in Section 2, we believe it would not improve the readability or accessibility of the paper. However, if the editor prefers a more mathematical description of each strategy, we are willing to include the equations. Since we are not conducting a mathematical analysis of the strategies, we have opted for a verbal description supported by code implementations.

- Reviewer 4 (comment 4) asked to add descriptive statistics. We do not believe the addition of descriptive statistics will add anything useful to the manuscript. We argue that the usefulness of descriptive statistics for time series analysis is limited.  We have provided the raw data, and we have provided time plots to help visualize the data. Each data set is described in Section 3.

- Reviewer 4 (comment 5) recommended including three papers from the literature. However, these papers are not relevant to our current study. The suggested papers focus on forecasting COVID-19 variables, which is not the focus of our paper.

@@

_Comment 1: The topic is very interesting and very timely, in the wake of the Covid -19 related disruptions in supply chains, changes in consumer preferences, and inflationary environment. The time series patterns and causal relationships that were seen before the pandemic did not apply and whether the patterns observed before the pandemic will continue to be relevant remains to be evaluated on a case by case basis._

**Response:** We are delighted to hear that you appreciate the timeliness and importance of the paper. We agree that the topic is particularly relevant given the COVID-19-related disruptions in supply chains and various other areas. We also agree that evaluating whether pre-pandemic time series patterns and causal relationships remain applicable is a critical challenge that requires careful case-by-case analysis. That is why, in this paper, we propose various strategies rather than a single approach to address different situations and cases effectively.

We have added the following paragraph in the revised manuscript to address this comment:

\textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis. Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}

@@

**Response:** Thank you for your comment. The idea for this paper emerged from our close contact with practitioners, who highlighted the challenges they face in forecasting time series data affected by COVID-19, particularly in handling observations impacted by the disruption. Following these discussions, we carried out a search to find out if any existing papers addressed this issue, but we could not find any published research in this area. To our knowledge, this is the first study to describe and compare general strategies for forecasting interrupted time series, such as COVID-19-affected data.

_Comment 3: The paper is well written, but in terms of contributions, it falls short of the expectations for an original research paper. Improvement avenues are proposing a viable new method that has advantages over these existing methods, or a large-scale empirical research with carefully curated set of actual data representing different interruption patterns from different domains._

Response: Thank you for your feedback. We appreciate your comments on the writing quality and your recommendations for improving the paper's contributions. We believe that the originality of our research comes from addressing the actual challenges that many forecasters and modellers experience when their time series data is disturbed by recommending multiple viable and reliable strategies suitable for different cases. We recognise that presenting a new technique may be an exciting direction, but it is not the aim of this paper. So we have included it as a potential topic for future research in this area. Also, we included two empirical datasets from two different application areas in the paper. Moreover, the whole paper is reproducible, and the datasets and R codes are accessible on GitHub, allowing strategies to be tailored to any dataset in any application.

\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events. Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data. Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}

@@

**Response:** We believe that our research tackles fundamental challenges that many forecasters and modellers face when working with time series data that has been disrupted by events such as COVID-19, strikes, etc. This challenge was highlighted throughout our conversations with academics and practitioners in the field, who emphasised the issues they face when forecasting such time series data. Therefore, we think addressing this practical challenge is where our work stands out. Further, we have not found any other papers addressing this issue. Consequently, we feel that we are addressing an important and widespread problem.

Additionally, our work may open avenues for further research and reflection in this area. Also, we made sure that the entire work was reproducible, with datasets and R codes available on GitHub, allowing other academics and practitioners to tailor the strategies to their own needs.

_Comment 2. The patterns of interrupted events may be very different, which makes the suitable approach varies. In this case, it may be better to focus on a certain event (e.g., COVID-19 pandemic) and make comprehensive comparisons._

**Response:** We agree that the patterns of interrupted events may vary, which would affect the suitability of proposed approaches. That is why we have proposed different strategies to consider that might be appropriate in different cases, each with its own pros and cons. Focusing on one event would greatly limit the scope of the paper, making it far less useful.

@@

In future work, a large scale empirical study covering several types of disruption could be conducted to provide deeper insights and any possible link between strategies and types of disruption events. We have included the following paragraph in the conclusion section of the revised manuscript:

\textcolor{teal}Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.}

@@

_Comment 3. The authors provide some visualizations on the forecasting. This should be appreciated since it is quite straightforward. However, the authors may not expect the readers to judge the optimal strategy from these figures. Instead, the authors should consider some out-of-sample validation and may introduce some metrics or measures to evaluate the performance different strategies._

Response: The visualizations are intended to help the reader understand the effect and intent of different strategies. In determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used. We think the visualizations help the reader do that. Also, it does not always make sense to do a quantitative comparison of forecast accuracy. We have attempted to address this issue by adding the following paragraph in the discussion section.

\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used. This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy. The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption. For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations. In both cases, it does not make sense to evaluate the performance of the method during the disruption. Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}

@@

_Comment 4. If the authors focus on the interrupted time series for a certain event, some exogeneous variables are encouraged for potential better performance._

**Response:** Thank you for your suggestion. We agree that incorporating exogenous variables related to specific events could enhance the performance of forecasting models for interrupted time series. We have acknowledged this by adding the following paragraph to the revised manuscript:

\textcolor{teal}{We should also note that when forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}

@@

_General comment: The manuscript deals with forecasting interrupted time series for forecasting teams, especially in light of emergency events like COVID-19. The manuscript investigates several strategies for dealing with interruptions in time series forecasting, and their advantages and disadvantages. The authors evaluated two datasets to understand the effectiveness of these strategies. The manuscripts deal with important areas and insight into the strategies._

A few suggestions need to be included such as:

_Comment 1: Different strategies are discussed in section 2 in a few words without giving a proper introduction of the strategies. It makes the reading of the manuscript difficult. The authors are required to give practical examples of the strategies to make it understandable to read._

**Response:** Thank you. We are not clear what is being requested here. We describe the strategies in Section 2, and we provide fully reproducible code showing how they are implemented. We also give practical examples of the strategies in Section 3. Please explain what more is required to be included, and we are willing to do that.

@@

_Comment 2. No mathematical support is provided as to how the strategies work. And what are the practical implementation of the strategies? Authors are required to provide detailed insight into the strategies i.e. which datasets these strategies are applied?_

**Response:** Thank you for the comment. We have carefully thought through this request. While we could add equations in Section 2 for each strategy, we do not think it would add to the readability or accessibility of the paper. We are not undertaking a mathematical analysis of the strategies, so we have opted for a verbal description accompanied by code that implements each strategy. However, if the editor would prefer us to add equations providing a more mathematical description of each strategy, we can do so. To respond to other remarks:

- The practical implementation of the strategies is provided in Section 3 and in the supplementary code.

- The data sets are described in Section 3.

- Insights are provided in Sections 3 and 4.

_Comment 3. The figures presented in the manuscripts require detailed explanation. Authors are required to explain the given figures in detail one by one._

**Response:** Thank you for your feedback. We have provide a detailed explanation of each figure in the manuscript, as well as in the captions of each figure, discussing them one by one to ensure clarity.

@@

_Comment 4. Two datasets are used and the strategies are applied to them. However, the descriptive statistics of these data sets missing.. The authors are required to explain the datasets._

**Response:** Thank you for this comment. We do not believe the addition of descriptive statistics will add anything useful to the manuscript. We argue that the usefulness of descriptive statistics for time series analysis is limited. First, they cannot highlight patterns in the data, identify unusual observations, describe changes over time, or help to understand relationships between variables. Second, they could be misleading, since similar statistics can stem from very different datasets (please see Anscombe’s Quartet).

We have provided the raw data, and we have provided time plots to help visualize the data. Each data set is described in Section 3.

### 2024-05-16 fits.qmd
finalising response letter, and revise the manuscript.

Time series forecasting models use historical data to estimate future values [@fildes2008forecasting] based on consistent patterns observed in the past. However, time series are sometimes disrupted by unusual events that jeopardise the regularity of the time series pattern. This can be defined as an *interrupted time series*. Time series data may be disrupted by a variety of reasons, ranging from temporary policy changes to natural disasters. Outbreaks, epidemics, and pandemics, for example, may have a considerable impact on population behavior, causing discrepancies and interruptions of data in a variety of industries such as healthcare, pharmaceuticals, transportation, retail, tourism, and traffic management. Similarly, supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps. Transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points. Furthermore, cybersecurity breaches data quality by possibly skewing readings in datasets relating to financial transactions, network traffic, and user activity, complicating time series analysis.

@@

In this section, we describe several possible strategies to handle interruptions when forecasting time series data.

@@

@@

This has the advantage of retaining the memory of the past, and so the seasonal patterns and other dynamics will be retained. This allows the change period to be effectively modelled, provided the intervention variables are chosen well. However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.

@@

@@

Future research might focus on enhancing probabilistic forecasting methods that are designed specifically for interrupted time series. Large-scale empirical research evaluating the effectiveness of different strategies in real-world settings might also provide useful insights. Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.

@@

>>>>

Time series forecasting models use historical data to estimate future values [@fildes2008forecasting] based on consistent patterns observed in the past. However, time series are sometimes disrupted by unusual events that jeopardise the regularity of the time series pattern. This can be defined as an *interrupted time series*. Time series data may be disrupted by a variety of reasons, ranging from temporary policy changes to natural disasters. Outbreaks, epidemics, and pandemics, for example, may have a considerable impact on population behavior, causing discrepancies and interruptions of data in a variety of industries such as healthcare, pharmaceuticals, transportation, retail, tourism, and traffic management. Similarly, transportation-related data may be interrupted due to equipment failures, such as malfunctions or breakdowns in vehicles, aircraft, or rail systems, resulting in missing or inaccurate data points. Furthermore, cybersecurity breaches data quality by possibly skewing readings in datasets relating to financial transactions, network traffic, and user activity, complicating time series analysis. Supply chain disruptions induced by factors such as strikes or shortages cause irregularities in sales, inventory levels, production, and delivery, often resulting in data gaps. \textcolor{teal}{The war in Ukraine caused disruptions in sunflower oil and wheat supply chains, which were most likely driven by social media and people's recent memories of COVID-19 toilet paper shortage. A power outage in Eastern Germany halted production at a Tesla facility and cut off electricity to a large retailer distribution centre, disrupting supplies to around 500 stores in and around Berlin for several weeks. Recent strikes in various forms of transport, including air and rail in Germany, France, and the United Kingdom, have had an impact on passenger flow and retail sales at airports and train stations. These issues highlight the significance of robust forecasting systems in addressing such disruptions effectively.}

@@

In this section, we describe several possible strategies to handle interruptions when forecasting time series data. \textcolor{teal}{We note that the assessment of whether pre-pandemic time series patterns and underlying relationships remain applicable is a critical issue that requires careful case-by-case analysis. Therefore, in this paper, we propose multiple strategies rather than a single approach to effectively address different situations and cases.}

@@

\textcolor{teal}{The disadvantage of smoothing and other highly adaptive techniques is that they need time to adapt after a disturbance and can eventually end up chasing noise. The latter is an intriguing issue since individuals often prefer adaptability while failing to recognise that it does not necessarily improve forecasting, ignoring the presence of noise in the data.}

@@

This has the advantage of retaining the memory of the past, and so the seasonal patterns and other dynamics will be retained. This allows the change period to be effectively modelled, provided the intervention variables are chosen well. \textcolor{teal}{Intervention models are useful because they enable the forecasting of similar future disruptions, but distinguishing between a disruption and a general driving factor, such as a promotion or a seasonal storm, might become subjective.} However, the model will assume that the post-interruption period will be similar to the pre-interruption period, and so the prediction intervals may be too narrow, especially if there is a lasting effect beyond the end of the interruption.

@@

\textcolor{teal}{We should also note that when forecasting interrupted time series data, exogenous variables associated with specific events should be identified and incorporated, since they may improve forecasting model performance.}

@@

\textcolor{teal}{Moreover, in determining which strategy to apply in each case, it is important to think through what is being assumed, and what data are being used. This is a qualitative and subjective process, and cannot simply be tackled with a quantitative comparison of forecast accuracy. The strategies we discuss have different ways of handling the disrupted period, and many of them do not seek to provide accurate forecasts during the disruption. For example, the "set to missing" strategy ignores the disrupted period, while the "estimate what might have been" strategy replaces the observations during the disruption with interpolations. In both cases, it does not make sense to evaluate the performance of the method during the disruption. Instead, the analyst needs to decide what approach makes the most sense for the application at hand.}

\textcolor{teal}{This study proposes multiple viable strategies for forecasting time series data interrupted by different type of events. Future research could focus on exploring new strategies for handling such disrupted time series, or enhancing probabilistic forecasting methods that are designed specifically for such data. Large-scale empirical research on further exploring which strategy may work best in different scenarios also provide useful insights.} Comparative studies that assess the accuracy, robustness, and computing efficiency of different techniques across several datasets and disruption situations can help discover the most effective approaches for a range of practical applications.

@@

# Acknowledgment {.unnumbered}

We would like to thank Dr. Stephan Kolassa for sharing his thoughts and providing valuable feedback on the first version of our paper. His voluntary feedback and recommendations have helped us to refine and improve our work.
